---
title: "Metropolis-Hastings for Beta distribution"
output: pdf_document
---


We write a Metropolis-Hastings algoritm to generate X~beta(6,4).
We select a r.v. from unif(0,1) as a starting value and a Beta jumping distribution  
where c is a parameter in our function. Later on we can test MH algoritm with differnt c's
10 percent of iterations will be treated as burn-in
```{r}
MHalg<-function(iter,c){
  
  phi<-  c()
  phi[1]<-runif(1,0,1)#generate a starting value from unif(0,1)

  new_phi<- function(phi) {
    proposal_phi <- rbeta(1, shape1= c*phi , shape2= c*(1-phi) ) #giving jumping dist.
    # ratio of accepting proposal_phi
    r <- (dbeta(proposal_phi, shape1 = 6, shape2 = 4)/
          dbeta(proposal_phi, shape1= c*phi , shape2=c*(1-phi)))/
         (dbeta(phi, shape1 = 6, shape2 = 4)/
          dbeta(phi, shape1= c*proposal_phi, shape2=c*(1-proposal_phi)))
  
    if (runif(1) <= r)
      return(proposal_phi)   
    else  
      return(phi)
  }

  for (i in 1:iter)
    phi[i+1]= new_phi(phi[i])
  phi=phi[-(iter/10)] #truncate burn-in
  
}
```
\newpage
For comparion, generate a histogram with random variables~beta(6,4)

```{r}
a1=rbeta(10000,6,4)
hist(a1,freq=FALSE,main="beta(6,4) histogram")
```
\newpage
We first start a simulation with c=1, and 10000 draws. Perform KS test to determine whether our generated random variables are numerically
approximately follow Beta(6,4)
```{r}
draw=MHalg(10000,1)
par(mfrow=c(1,3)) 
plot(draw); acf(draw); hist(draw,freq=FALSE) 
title("C=1", outer=TRUE)

ks.test(draw,a1) 
```
\newpage
Now we simulate with c=0.1, 2.5, 10.
```{r}
draw=MHalg(10000,0.1)
par(mfrow=c(1,3)) 
plot(draw); acf(draw); hist(draw,freq=FALSE)  
ks.test(draw,a1) 
title("c=0.1", outer=TRUE)
```
\newpage
```{r}
draw=MHalg(10000,2.5)
par(mfrow=c(1,3)) 
plot(draw); acf(draw); hist(draw,freq=FALSE)  
ks.test(draw,a1) 
title("c=2.5", outer=TRUE)
```
\newpage
```{r}
draw=MHalg(10000,10)
par(mfrow=c(1,3)) 
plot(draw); acf(draw); hist(draw,freq=FALSE)  
ks.test(draw,a1) 
title("c=10", outer=TRUE)
```
Comparing histograms and Kolmogorov–Smirnov statistic, c=10 is the most effective.

Now we lower the iteration to 3000. The result is already very close to beta(6,4)
```{r}
draw=MHalg(3000,10)
par(mfrow=c(1,3)) 
plot(draw); acf(draw); hist(draw,freq=FALSE)  
a1=rbeta(3000,6,4)
ks.test(draw,a1) 
```






